---
title: "Proposal assessment"
author: "Andrew Irwin"
date: "2022-11-01"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
library(janitor)
```

```{r echo=FALSE}
grade <- tribble(
  ~ Component, ~ Score, ~ Maximum,
  "Questions", NA, 2,
  "Data provided and read into R", NA, 1,
  "Data description", NA, 2,
  "Visualization ideas", NA, 2,
  "Team planning", NA, 2,
  "References", NA, 1
  )
grade$Score = c(2, 1, 2, 2, 2, 1)
grade %>% 
  adorn_totals() %>%
  kable() %>%
  kable_styling(full_width = FALSE)
```

## Feedback

I will revise this section to provide feedback on your proposal. You can think of these sentences as a rubric that distinguishes between excellent, good, and insufficient proposals.

Your questions were
easy to understand and of appropriate complexity.

You 
provided your data in the `data` folder and showed how it could be read with R.

You described the source, observations and variables in your data well enough so that I could make a connection between your data and your questions.

You described appropriate visualizations for your data and questions.

You have a clear plan for the work of each team mate.

You provided a reference so I know where your data came from.

## Comments

If there are additional comments I'll add them here. If not, this paragraph will remain unchanged.

You can use categorical variables (e.g., smoking status, marital status) as predictors in a linear model. There are some pitfalls, particularly when interpreting the results, but go ahead and explore the methods.

